On the use of decision trees to identify the local extrema in parallel global optimization algorithms

### 2023-01-30 2023/PAVT/ParallelDecisionTree.tex

The Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods. Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{NelderMead,Himmelblau75}.

@@

Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (GSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented. The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard) of dimensionalities 2, 3, 4, and 5 from \cite{Sergeyev2006}. The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.  In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of speedup in time relative to the sequential run are compared. The  parallelization was performed using MPI technology, the run was performed on 8 processes.

>>>>

The Hooke--Jeeves pattern search algorithm belongs to the class of zero-order methods. Its computation rules are a  combination of the investigating search (to select the direction) and search in the direction selected  \cite{Himmelblau75}.

@@

Below, the results of comparing two parallel algorithms --- asynchronous global search algorithm  (AGSA) and its modification using the decision tree to find the attractor areas of the local minima  (Decision Tree) are presented. The numerical comparison was performed on two classes of the GKLS  functions (Simple and Hard from \cite{fio_bib19}) of dimensionalities 2, 3, 4, and 5. The stop  criterion for the algorithms was the falling of the next trial points into the $\varepsilon$-nearness of  true global minimum.  In Table \ref{tab:2}, the averaged numbers of iterations executed by the algorithms when solving the  problems and the values of time speedup relative to the sequential run are compared. The parallelization was performed using MPI technology, the run was performed on 8 processes.
